---
title: "Numerical~Cat template"
author: "Dr K"
date: "November 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# t.test

This is a template to serve as an example (cut and paste, change variable names)  
we use the ucdavis1 dataset from the tigerstats package.

```{r include=FALSE}
require(tidyverse)
require(tigerstats)
data(ucdavis1)
davisdata <- na.omit(ucdavis1)
OurData <- chickwts
```


### Plot boxplots

```{r}
plot(momheight~Sex,data=davisdata)
```

## Now with ggplot - first select the basic data

```{r}
basicCN <- ggplot(ucdavis1,aes(x=Sex,y=dadheight))
```
## Now add in boxplot

```{r}
basicCN + geom_boxplot()
```
  
  
### get fancy  

```{r}
basicCN + geom_boxplot(notch=TRUE,outlier.color = "red") + stat_boxplot(geom ='errorbar') + labs(title="Fancy notch boxplot \n with 2 lines")
```

# Numerical results

```{r}
favstats(momheight~Sex,data=ucdavis1)
```

# Inferential  
  
```{r}
t.test(momheight~Sex,data=ucdavis1)
```
# One Way Anova
  
  

### Plot boxplots

```{r}
plot(alcohol~Seat,data=davisdata)
```
   
#### Using ggplot

```{r}
basicCN <- ggplot(davisdata,aes(x=Seat,y=alcohol))

basicCN + geom_boxplot()
```
  
# Numerical results

```{r}
favstats(alcohol~Seat,data=ucdavis1)
```
# Inferential

```{r}
model1 <- aov(alcohol~Seat,data=davisdata)
summary.aov(model1)
```

## When significant follow up with TukeyHSD  

```{r}
plot(TukeyHSD(model1))
```

## Another example  

```{r}
plot(weight~feed,data=OurData)
```
  
  
### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(weight~feed,data=OurData)
```

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(weight~feed,data=OurData)
summary.aov(model1)
```
```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 8.7, 2, 2))
plot(TukeyHSD(model1),las=2)
```
## Repeated Measures Anova  

### This approach is going to use the LMER package and treat the subjects as random effects


```{r include=FALSE}
require(lme4)
require(lmerTest)
mydata <- sleepstudy
str(mydata)
```
## lets fit a linear model explaining Reaction as a function Days

```{r}
basicNN <- ggplot(mydata,aes(y=Reaction,x=Days))
```
##Now add in scatterplot

```{r}
basicNN + geom_point() + geom_smooth(method="lm")
```


```{r}
model.1 <- lm(Reaction~Days,data=mydata)
summary.lm(model.1)
```
  
## Now a random effects model 
```{r}
 fm1 <- lmer(Reaction ~ Days + (Days | Subject), mydata)
summary(fm1)
```
