---
title: "WUP_04"
author: "Shaun ikeuchi"
date: "2/28/24"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)

```

# Introduction
```{r}
data <- read.csv(file="GSScsv.csv")  

```
<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  


# Methods 
 
The numerical variable is the average score on the WORDSUM test and the categorical variable is the level of schooling Degree. 
Thus we will use the helper file numerical~categorical.


<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results
```{r}
basicCN <- ggplot(data,aes(x=WORDSUM,y=DEGREE))

basicCN + geom_boxplot()
```



 According to the boxplot, there shows difference in median values between every degree with LT high school being the lowest with a value of 8 compared to average score on the WORDSUM test in the graduate with the median as 11.
<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Numerical Descriptive Results
```{r}
favstats(WORDSUM~DEGREE,data=data)
```
 


 There is a deviation in the mean based on degree with the lowest being LT High school with an average score of 7.75 compared to the Graduate with the average score to be 10.095.
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

## Inferential Results
```{r}
model1 <- aov(WORDSUM~DEGREE,data=data)
summary.aov(model1)
```
Since the p-value is 0.000236 which is less than 0.05 there is evidence that this is statistically significant.
<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion
Our question is whether or not there is a difference between average score on the WORDSUM test and the level of scoring. According to the boxplot, the median values differ as well as the minimum and maximum values differ significantly with the degree with LT high school being the lowest with a value of 8 compared to average score on the WORDSUM test in the graduate with the median of 11. In addition, there is a deviation in mean with the lowest being LT High school with an average score of 7.75 compared to the Graduate with the average score to be 10.095. Furthermore, the p-value of 0.00236 is less than 0.05 providing further evidence that this is statisically significant. Based on this data, we can reject the NULL hypothesis and therefore demonstrate that there is a statisically significant difference between degree and WORDSUM test score.
<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
